---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/ropenscilabs/pkginspector.svg?branch=master)](https://travis-ci.org/ropenscilabs/pkginspector)
[![codecov](https://codecov.io/gh/ropenscilabs/pkginspector/branch/master/graph/badge.svg)](https://codecov.io/gh/ropenscilabs/pkginspector)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of pkginspector is to inspect and better understand the internal structure of an R package, with focus on facilitating package reviews and [**rOpenSci** onboarding](https://github.com/ropensci/onboarding).

## Installation

You can install pkginspector from GitHub with:

```R
# install.packages("devtools")
devtools::install_github("ropenscilabs/pkginspector")
```

## Example

The inspected package must be stored locally and installed.

```{r rev_fn_summary}
library(pkginspector)

path <- pkginspector_example("viridisLite")
knitr::kable(rev_fn_summary(path))
```

```{r rev_args}
rev_args(path)

corrplot::corrplot(rev_args(path)$arg_map, method = 'square', cl.pos = "n")
```

## Information

* [Getting help](SUPPORT.md).
* [Contributing](CONTRIBUTING.md).
* [Contributor Code of Conduct](CODE_OF_CONDUCT.md).

